Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-07-17 manuscript/revision2/response_to_reviews/response_to_reviews.qmd

1.  Initialize covariance matrices of interest
2.  Initialize lists to store the most likely additional matrix and corresponding maximum likelihood
3.  Obtain model likelihood for a model fit with each matrix
4.  Remove the matrix with the greatest likelihood from the array and store it
5.  Repeat the process of fitting mash models, finding the most likely, pair, trio, etc. (using the most likely matrices from the preceding iteration) until the likelihood ratio test is no longer significant

![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the green-up date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. (A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. (B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. (D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. (E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure1.png)

![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering date phenotype using only GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. (A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. (B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. (D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. (E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure2.png)

![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the greenup date phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. (A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted. (B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. (D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted for computational considerations. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. (E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure3.png)

![**Log-likelihood at each step in a greedy algorithm implementation of model selection for mash for the flowering phenotype using canonical, data driven, and GxWeather covariance matrices.** The matrix that resulted in the maximum likelihood model of covariance in summary statistics among sites is located at the base of the y-axis; matrices selected in the ensuing step are placed in ascending order along the y-axis. (A) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent variants that were significant in at least one context when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted. (B) and (C) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively. (D) The maximum likelihood path of covariance matrices in the greedy algorithm for the set of independent, randomly selected variants when analyzing all individuals. The dashed black line represents the step in the algorithm where addition of the next matrix did not significantly increase the model likelihood, likelihood-ratio test p-value \> 0.05. Once this condition was met the greedy algorithm was halted for computational considerations. \* is used to denote those matrices that are identified by the greedy algorithm in the set of significantly associated variants and the randomly associated variants. (E) and (F) show the maximum likelihood paths when the individuals analyzed are only from the Gulf and Midwest subpopulations, respectively.](images/RR_Figure4.png)

@@

We now use the narrow-sense heritability as the diagonal for each GxWeather matrix, and clarify this in the Supplementary Methods (SI Appendix, Section S1, first and eighth paragraph - first and last paragraph on page two).

> ::: {.callout-warning appearance="minimal"}
> \- Antagonistic Pleiotropy: Maybe I’m missing the idea of this analysis, but it doesn’t seem to me that this strategy for partitioning loci between antagonistic pleiotropy and differential sensitivity would have equal power to detect each class. It seems that antagonistic pleiotropy should be a subset of differential sensitivity because the latter is defined as simply a change in magnitude. But even if it’s restricted to “different magnitude but same sign”, the practical definition here seems to be that antagonistic pleiotropy is detected when the signs are different and both lfsrs \< threshold (0.05?), while differential sensitivity additionally requires a threshold on the difference in effect sizes (0.4) with no justification for why this size was chosen. This additional criterion will necessarily make the rate of detection different. Even without this, with differential sensitivity one effect size must be much larger than the other in absolute value (which isn’t the case with antagonistic pleiotropy), so differential sensitivity loci will be more likely to pass the lfsr loci than antagonistic pleiotropy loci. Note that previous literature here tried to separate antagonistic pleiotropy from conditional neutrality (ie where in one location the effect was zero), while here the contrast is with differential sensitivity where effects are non-zero in all locations. I’m not aware of previous discussions in the literature of trying to differentiate these two specific classes of GxE variants and not really sure theoretically what the importance would be.
> :::

We believe that the use of the lfsr to detect effects with rank-changing GxE is a key advance in our manuscript, and so we have made changes to the Introduction (lines 169-171), Results (lines 319-326), and Materials & Methods (lines 957-997; lines 1055-1056) to explain and motivate this statistical change.

Our expectation is that in nature, effects will always differ. Statistically, this difference will not always be detectable. Thus, rather than ask "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?

We think that if evolutionary geneticists are interested in antagonistic pleiotropy and rank-changing GxE, then they should use the lfsr to measure their confidence in the sign of the effect, rather than using statistical tests where the null hypothesis is that the effect is different than zero. However, if we use the lfsr, comparisons between antagonistic pleiotropy and conditional neutrality are no longer sensible, as we are no longer doing a statistical test that can robustly detect conditional neutrality (just as the FDR does not robustly detect antagonistic pleiotropy). This is justified because detection of antagonistic pleiotropy is more important than detection of conditional neutrality for models of local adaptation.

It's true that the threshold for differential sensitivity is arbitrary. We have reworded the manuscript to stress that there is equal power to detect effects of different sign as there are effects of the same sign (lines 994-997; lines 1055-1056). Another class of effects we have in our analysis that differ from classic antagonistic pleiotropy & conditional neutrality comparisons are effects which are not distinguishable by sign or magnitude. It is not accurate that differentially sensitive loci will be more likely to pass the lfsr, because the loci are first selected by significant lfsr, then divided further into effects that can & cannot be distinguished by magnitude. Thus effects that are significant, but have small magnitude in both conditions will be labeled as not being distinguishable - essentially, these effects do not have detectable GxE. Effects without GxE may be even more common and should also be quantified. We propose that, if our intent is to detect sign-changing GxE, then we should use an unbiased statistical test to look at loci with and without sign-changing GxE, the lfsr.

::: {.center data-latex="1"}
Best,
:::

::: {.flushright data-latex="1"}
Alice MacQueen and Tom Juenger
:::

### 2024-07-17 manuscript/revision2/response_to_reviews/response_to_reviews.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tikz}{}{\usepackage{tikz}}
\makeatother
        \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
          \node[shape=circle,draw,inner sep=1pt] (char) {{\scriptsize#1}};}}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\begin{flushright}
The University of Texas at Austin\\
Department of Integrative Biology\\
2415 Speedway\\
Austin, TX 78712

\end{flushright}

~

\begin{flushright}
July 17th, 2024

\end{flushright}

~

\begin{flushleft}
Editors\\
Proceedings of the National Academy of Sciences\\

Subject: Manuscript revision

\end{flushleft}

~

@@

Dear Editors,

Thank you for considering this work for publication in \emph{PNAS}. Our
manuscript, ``Diverse Genotype-by-Weather Interactions in Switchgrass''
has undergone substantial revision from our original submission in 2021.
This timeframe is long - Dr.~MacQueen took a new position in 2022. They
continued working with the group as we developed a new algorithm to
address major reviewer concerns and significantly altered the
manuscript. Because of the major alterations over this time frame, the
response to reviewers does not have track changes for each point, though
we do point to line numbers for the associated revisions in our
response.

We were pleased by the useful feedback given by the reviewers of our
original submission and believe that the greedy mash algorithm we
implemented addresses the major statistical concern of reviewer 2, as
well as substantially improving our use and interpretation of covariance
matrices in \emph{mash}. We have addressed all the editorial issues
suggested by the two reviewers, and included in this letter are
point-by-point responses to these issues.

We suggest that the original editor, Editorial Board members, and
anonymous reviewers for this manuscript would be good choices to assess
this revision. In the event that is not practicable, Douglas Schemske,
Stephen P. Long, and Qifa Zhang would be good Editorial Board members
for this manuscript. All would be able to provide strong insights into
mapping the genetic basis of adaptive and ecologically relevant traits
in a field experimental context. Johanna Schmitt, Trudy Mackay, and
Detlef Weigel would be excellent editors for this manuscript.
Dr.~Schmitt has done extensive work in \emph{Arabidopsis thaliana} and
other plant species unravelling the genetic mechanisms involved in
responses to the natural environment. Dr.~Mackay has extensive
experience mapping the genetics of complex traits and
genotype-by-environment interactions. Dr.~Weigel has done extensive work
understanding genetic diversity in \emph{A. thaliana} including
extensive work on floral induction cues.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]

Reviewer \#1:\\
\strut \\
Suitable Quality?: No\\
Sufficient General Interest?: No\\
Conclusions Justified?: Yes\\
Clearly Written?: No\\
Procedures Described?: Yes\\
\strut \\
Comments on Significance Statement:\\
\strut \\
Conclusions are very specific for the crop/population and traits.
General conclusions are lacking.\\
\strut \\
Comments:\\
\strut \\
The paper by MacQueen et al.~presents results of a study on two
switchgrass populations grown in 8 garden experiments across a rane of
latitudes. Traits are green-up date and flowering time. The data show
strong GxE, effects of QTL changing in magnitude and sign. Results are
partly validated in an independent segregating population.\\
\strut \\
The topic is timely and highly relevant. A better understanding of GxE
will be valuable for plant genetics, evolution and breeding. The data
are complex and the authors have succeeded in making them accessible to
the reader.\\
\strut \\
\strut \\
Relevant literature\\
The introduction and discussion are focused on findings in Arabidopsis
and switchgrass. An excellent study from Arabidopsis with high relevance
for this paper is\\
Fournier-Level et al.~2016
\url{https://doi.org/10.1073/pnas.1517456113}\\
\strut \\
Major findings from other crop species have been ignored. To mention
just two references:\\
\strut \\
Bustos-Korts et al.~2019 The Plant Journal
\url{https://doi.org/10.1111/tpj.14414}\\
Millet et al.~2016 Plant Physiology
\url{https://doi.org/10.1104/pp.16.00621} and\\
There are many papers on crop modelling (eg work by Mark Cooper and
others in maize) and the authors might want to consider linking their
results to crop modelling to demonstrate the relevance of the results of
this study

\end{tcolorbox}
\end{quote}

@@

We now add these citations (lines 47-50; lines 58-61). We regret we
can't comprehensively reference the impressive crop modeling literature.
Our goal was to reference work on natural populations that focused the
genetics on phenology and/or local adaptation, and to touch on how these
ideas have interacted with crop improvement. This was because our work
uses a natural population of switchgrass with potential relevance for
breeding efforts in this species.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]

Results\\
\strut \\
There is a large overlap of data and findings with reference 32 Lovell
et al.~2021 Nature. It has not become clear which additional insights
can be gained from this study. One reason might be that objectives of
the study have remained rather vague as well as the generic conclusions
that can be drawn from the results.

\end{tcolorbox}
\end{quote}

While we do rely on the same genotypic data and common gardens used in
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021}, and make use of the population structure findings of
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021} to divide our genotypes into subpopulations for GWAS, the previous
work considered fitness proxies and this work considers phenological
traits. This work has two additional, key advances over the previous
work: it can assign loci genome-wide to multiple kinds of GxE and
GxWeather, and has an unbiased statistical method to identify loci with
and without rank-changing GxE. We now summarize the findings from
\href{https://doi.org/10.1038/s41586-020-03127-1}{Lovell \emph{et al.,}
2021} that we use as a springboard for this work on lines 136-145, and
outline the additional advances of this study on lines 156-163, 169-171,
and 883-893.

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]

In the introduction the aim is given as „we test if these populations
differ in their phenological adaptation and hence their phenological
GxE''. Given prior knowledge such as the origin of the populations and
earlier results (Lovell et al.) what is expected?

\end{tcolorbox}
\end{quote}

@@

We now edit the Introduction to more clearly articulate the aims of this
manuscript. Our key expectation is that different genetic subpopulations
and genomic regions have likely evolved distinct patterns of GxE (lines
55-58; lines 219-223). Thus, we aim to identify the kinds of GxE present
for each subpopulation and each trait. (lines 129-131).

\begin{quote}
\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, arc=.35mm, bottomrule=.15mm, opacityback=0, colback=white]

The results for these two populations did not show a general pattern
leading to conclusions on the environmental cues. Associations trait/cue
changed across populations, traits, gardens, analysis etc. , it was
difficult to see the big picture.

\end{tcolorbox}
\end{quote}
